The Ontological Modal Collapse \\ as a Collapse of the Square of Opposition

### 2014-08-31 Papers/Universalis/ModalCollapse.tex
Remove wrong claims

A crucial step of the ontological argument is the claim that if God's existence is possible, then it is necessary. This is Lemma L2 in G\"odel's proof. In the modal square of opposition (Fig. \ref{ToDo}), this is an unusual situation in which the \textbf{I} corner implies and entails the \textbf{A} corner, for the particular case when $\phi$ is $\exq x G(x)$. G\"odel's proof shows that his axioms are strong enough to invert the direction of entailment for the particular $\phi$ at issue. The question is whether the axioms are then not too strong, also allowing the inverted entailment for arbitrary $\phi$. That's essentially the question asked by Sobel \cite{Sobel1987}; and his proof of the modal collapse (MC) answers it affirmatively. The modal collapse is a collapse of the modal square: it shows that assuming G\"odel's axioms, the subcontraries entail the contraries.

@@

ToDo: the following should always hold (check this in Isabelle!): 
$$
\allq \phi [\exq x G(x) \imp \phi] \imp [\phi \imp \nec \phi]
$$
$$
\allq \phi [\exq x G(x) \imp \phi] \imp [\pos \phi \imp \nec \phi]
$$

\begin{figure}[t]
\noindent \framebox[\columnwidth][r]{
\begin{minipage}{.94\columnwidth}\small
\begin{itemize}
\item[MC] Everything that is the case is so necessarily:
  $$\hol{\allq \phi [\phi \imp \nec \phi]}$$ 
\item[MC'] Everything that is possible is necessary:
  $$\hol{\allq \phi [\phi \imp \nec \phi]}$$ 
\item[MC''] Modalities collapse:
  $$\hol{\allq \phi [\phi \biimp \pos \phi \biimp \nec \phi]}$$
\end{itemize}
\end{minipage}
} \vskip-.5em
\caption{Modal Collapse \cite{todo:Sobel1987,SobelBook}.\label{fig:collapse}} 
\end{figure}

>>>>

A crucial step of the ontological argument is the claim that if God's existence is possible, then it is necessary. This is Lemma L2 in G\"odel's proof. In the modal square of opposition (Fig. \ref{ToDo}), this is an unusual situation in which the \textbf{I} corner implies and entails the \textbf{A} corner, for the particular case when $\phi$ is $\exq x G(x)$. G\"odel's proof shows that his axioms are strong enough to invert the direction of entailment for the particular $\phi$ at issue. The question is whether the axioms are then not too strong, also allowing the inverted entailment for arbitrary $\phi$. That is essentially the question asked by Sobel \cite{Sobel1987}; and his proof of the modal collapse (MC) answers it affirmatively. The modal collapse is a collapse of the modal square: it shows that, assuming G\"odel's axioms, the subcontraries entail the contraries.

@@

\begin{figure}[t]
\noindent \framebox[\columnwidth][r]{
\begin{minipage}{.94\columnwidth}\small
\begin{itemize}
\item[MC] Everything that is the case is so necessarily:
  $$\hol{\allq \phi [\phi \imp \nec \phi]}$$ 
\item[MC'] Everything that is possible is necessary:
  $$\hol{\allq \phi [\pos \phi \imp \nec \phi]}$$ 
\item[MC''] Modalities collapse:
  $$\hol{\allq \phi [\phi \biimp \pos \phi \biimp \nec \phi]}$$
\end{itemize}
\end{minipage}
} \vskip-.5em
\caption{Modal Collapse \cite{todo:Sobel1987,SobelBook}.\label{fig:collapse}} 
\end{figure}
